### 2017-11-13 paper/sections/1_introduction.tex
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

This last contribution is explored in
Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
and~\S\ref{sec:cache-management-using-domain-specific-knowledge}, where we try
a range of policies from different disciplines; but more importantly, in
Section~\S\ref{sec:relation-to-load-balancing}, we conclude that the collection
of policies we designed for ParSplice's cache management are very similar to
the policies used to load balance metadata in the Ceph file system (CephFS)
suggesting that there is potential for automatically adapting and generating
policies dynamically.

>>>>

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we designed for ParSplice's cache management
are very similar to the policies used to load balance metadata in the Ceph file
system (CephFS) suggesting that there is potential for automatically adapting
and generating policies dynamically.

### 2017-11-13 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The scalability experiment uses 1 splicer, 1 persistent database, 1 cache
process, and up to 2 workers. We scale up to 1024 tasks, which spans 32 nodes
and disable hyper-threading because we experience unacceptable variability in
performance. For the rest of the experiments, we use 4 nodes, 1 splicer, 1
persistent database, 1 cache process, 1 worker, and up to 512 tasks.  The
keyspace analysis that follows is for the cache on the worker node, which is
circled in Figure~\ref{fig:parsplice}.  The cache hierarchy is unmodified but
for the persistent database node, we replace BerkeleyDB on NFS with LevelDB on
Lustre. Original ParSplice experiments showed that BerkeleyDB's syncs caused
reads/writes to bottleneck on the persistent database. We also use Riak's
customized LevelDB\footnote{https://github.com/basho/leveldb} version, which
comes instrumented with its own set of performance counters.

@@

\subsubsection*{An active but small keyspace} The bars in
Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
worker tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when worker tasks generate segments for the same
state, which is computationally cheap (this motivates
Section~\S\ref{sec:cache-management-using-system-archiecture-knowledge}).

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
growth rate, temperature, and number of workers have a predictable effect on
the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the regimes and adapt the
storage system (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).

>>>>

The scalability experiment uses 1 splicer, 1 persistent database, 1 cache
process, and up to 2 workers. We scale up to 1024 tasks, which spans 32 nodes
and disable hyper-threading because we experience unacceptable variability in
performance. For the rest of the experiments, we use 8 nodes, 1 splicer, 1
persistent database, 1 cache process, 1 worker, and up to 256 tasks.  The
keyspace analysis that follows is for the cache on the worker node, which is
circled in Figure~\ref{fig:parsplice}.  The cache hierarchy is unmodified but
for the persistent database node, we replace BerkeleyDB on NFS with LevelDB on
Lustre. Original ParSplice experiments showed that BerkeleyDB's syncs caused
reads/writes to bottleneck on the persistent database. We also use Riak's
customized LevelDB\footnote{https://github.com/basho/leveldb} version, which
comes instrumented with its own set of performance counters.

@@

\subsubsection*{An active but small keyspace} The bars in
Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
worker tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when worker tasks generate segments for the same
state, which is computationally cheap (this motivates
Section~\S\ref{sec:dom-specific}).

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates Section~\S\ref{sec:dom-specific}).
The growth rate, temperature, and number of workers have a predictable effect
on the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) looks
like the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the regimes and adapt the
storage system (this motivates Section~\S\ref{sec:dom-specific}).

### 2017-11-13 paper/sections/3_methodology.tex
paper: addresses #11
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addressed #8
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%  \item background: CephFS implementation
To explore dynamic cache management policies that change during the run, we use
the data management language and policy engine presented
in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
and it was originally built on the CephFS so administrators could control file
system metadata data management policies. We now refer to Mantle as a policy
engine that supports our data management language.  The basic premise is that
load balancing policies can be expressed with a simple API consisting of
``when", ``where", and ``how much". ``when" controls how aggressive or
conservative the decisions are; ``where" lets the policy control how
distributed or concentrated the data should be; and ``how much" controls the
amount of data that should be sent.

@@

\subsection{Extracting Mantle as a Library}

@@

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
size although the other metrics proved to be valuable debugging tools. Using
Mantle time series metrics, we expose accesses to the cache as a list of
\texttt{timestamp, keyid} pairs. In
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}, we
explore a regime detection algorithm that uses this metric.

>>>>

%  \item background: CephFS implementation
To explore dynamic cache management policies that change during the run, we use
the data management language and policy engine presented
in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
and it was originally built on the CephFS so administrators could control file
system metadata data management policies. We now refer to Mantle as a policy
engine that supports our data management language.  The basic premise is that
data management policies can be expressed with a simple API consisting of
``when", ``where", and ``how much". ``when" controls how aggressive or
conservative the decisions are; ``where" lets the policy control how
distributed or concentrated the data should be; and ``how much" controls the
amount of data that should be sent.

@@

\subsection{Extracting Mantle as a Library}
\label{sec:extracting}

@@

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
size although the other metrics proved to be valuable debugging tools. Using
Mantle time series metrics, we expose accesses to the cache as a list of
\texttt{timestamp, keyid} pairs. In Section~\S\ref{sec:dom-specific}, we
explore a regime detection algorithm that uses this metric.

### 2017-11-13 paper/sections/4_dynamic-policies.tex
paper: hack on generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Cache Management Using System Architecture Knowledge}
\label{sec:cache-management-using-system-archiecture-knowledge}

@@

\section{Cache Management Using Domain-Specific Application Knowledge}
\label{sec:cache-management-using-domain-specific-knowledge}
\begin{figure}[t]
  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\

@@

\subsection{Regime Detection}

@@

\subsection{Trajectory Length}

\subsection{Cloud Techniques: Elastic Search}

What if we re-provision resources in response to events outside the
application's control, such as a slow Lustre.

\subsection{Future work}

How Much: cache policy from past, regime detection, Belady's Min

>>>>

%\subsection{Trajectory Length}
%
%\subsection{Cloud Techniques: Elastic Search}
%
%What if we re-provision resources in response to events outside the
%application's control, such as a slow Lustre.
%
%\subsection{Future work}
%
%How Much: cache policy from past, regime detection, Belady's Min

\section{Cache Management Using System Architecture Knowledge}
\label{sec:arch-specific}

@@

\section{Cache Management Using Domain-Specific Application Knowledge}
\label{sec:dom-specific}
\begin{figure}[t]
  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\

@@

\subsection{Regime Detection}
\label{sec:regime-detection}

@@

### 2017-11-13 paper/sections/5_file-systems.tex
paper: cleanup generality section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{General Data Management Policies}
\label{sec:relation-to-load-balancing}

@@

\subsection{Using File System Policies for ParSplice}
%- memory utilization/heuristics
From a high-level the ParSplice policy trims the cache if the cache reaches a
certain size {\it and} if it has already absorbed the initial burstiness of the
workload; the CephFS poicy, which is shown in Figure~\ref{src:lua-cephfs},
migrates load if the metadata load is higher than the average load {\it and}
the current load has been overloaded for more than two iterations.

\textbf{Condition for ``Overloaded"} (Fig.~\ref{src:lru-dyn}: Line 2;
Fig.~\ref{src:lua-cephfs}: Line 2) - these lines detect whether the node is
overloaded using the ``load" calculated in the load callback; while the load
calculations and thresholds are different, the actual logic is exactly the
same.  Recall that this decision is made locally because there is no global
scheduler or centralized intelligence.

\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
Fig~\ref{src:lua-cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded. The
underlying implementation saves the values to local disk.

\textbf{Condition that Switches Policy} (Fig.~\ref{src:lru-dyn}: Line 6;
Fig.~\ref{src:lua-cephfs}: Line 5) - these lines switch the policies using
information from previous decisions. ParSplice trims its cache once it eclipses
the ``absorb" threshold while CephFS allows balancing when overloaded for more
than two iterations. The persistent state is essential for both of these
policy-switching conditions.

@@

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is to keep caches on
both clients and servers. The caches in CephFS have the same performance and
utilization trade-off as ParSplice, where large caches improve performance at
the expense of a large memory footprint. High memory utilization is problematic
when a single metadata server services many clients with their own caches.
Using cache management policies has two benefits for load balancing: (1)
increases the capacity of a single metadata server, and (2) helps identify
which parts of the cache to keep local to a server. Applying the regime
detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
system access pattern has the potential to identify when to migrate keys and
how many keys to migrate.

% burstiness of creates then compile
To show how regime detection helps load balance file system metadata across a
cluster, consider the trace of metadata requests for compiling a large project
in CephFS shown in Figure~\ref{fig:compile-ops}. That trace shows the number of
file system metadata requests serviced by the metadata server when
uncompressing (\texttt{untar}), compiling (\texttt{make}), and deleting
(\texttt{rm}) the source code for the Linux kernel. If the system knows that
the job phases would progress from many creates, to many lookups, to many
deletes, then it could size its caches accordingly. For example, the file
system could cache none of the inodes from the \texttt{untar} phase and run
regime detection during the \texttt{make} phase and only cache the inodes that
are repeatedly used. This will be less than 20K inodes and could result in up
xMB in memory footprint savings without sacrificing performance.

%- strategy: cache all inodes on create
\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
\caption{ \label{fig:compile-ops}} \end{figure}

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in
Sections~\S\ref{sec:cache-management-using-system-archiecture-knowledge}
and~\ref{sec:cache-management-using-domain-specific-knowledge}. Administrators
can use the policies in ParSplice to automatically size and manage cache tiers
\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/}, caching on
object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain keyspace access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization and performance.

http://docs.ceph.com/docs/master/dev/mds_internals/data-structures/

>>>>

\section{General Data Management Policies}
\label{sec:scope}

@@

\subsection{Using File System Policies for ParSplice}

%- memory utilization/heuristics
From a high-level the cache management policy we designed in
Figure~\ref{src:lru-dyn} trims the cache if the cache reaches a certain size
{\it and} if it has already absorbed the initial burstiness of the workload.
Much of this implementation was inspired by the file system metadata load
balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
in~\cite{sevilla:sc15-mantle}. That policy migrates load if the metadata load
is higher than the average load in the cluster {\it and} the current load has
been overloaded for more than two iterations. The two policies have the
following in common:

\textbf{Condition for ``Overloaded"} (Fig.~\ref{src:lru-dyn}: Line 2;
Fig.~\ref{src:lua-cephfs}: Line 2) - these lines detect whether the node is
overloaded using the load calculated in the load callback (not shown). While
the calculations and thresholds are different, the way the loads are used is
exactly the same; the ParSplice policy flags the node as overloaded if the
cache reaches a certain size while the CephFS policy compares the load to other
nodes in the system.

\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
Fig~\ref{src:lua-cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded.

\textbf{Two-Part Policies} (Fig.~\ref{src:lru-dyn}: Line 6;
Fig.~\ref{src:lua-cephfs}: Line 5) - after determining that the node is
overloaded, these lines add an additional condition before the policy enters a
data management state.  ParSplice trims its cache once it eclipses the
``absorb" threshold while CephFS allows balancing when overloaded for more than
two iterations. The persistent state is essential for both of these
policy-switching conditions.

@@

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is caching. If
clients and servers maintain a consistent cache, the client can do operation
locally without contacting the metadata server.  The caches in CephFS have the
same performance and utilization trade-off as ParSplice, where large caches
improve performance at the expense of a higher memory footprint. High memory
utilization is problematic when a single metadata server maintains consistent
caches with many clients .  Using cache management policies from the previous
two sections has two benefits for load balancing: (1) increases the capacity of
a single metadata server, and (2) helps identify which parts of the cache to
keep local to a server. For example, applying the regime detection algorithm from
Section~\S\ref{sec:regime-detection} to a similar file system access pattern
has the potential to identify when to migrate keys and how many keys to
migrate.

% burstiness of creates then compile
To explore this example in more detail, consider the trace of metadata requests
for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
shows the number of file system metadata requests serviced by the metadata
server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
deleting (\texttt{rm}) the source code for the Linux kernel. If the system
knows that the job phases would progress from many creates, to many lookups, to
many deletes, then it could size its caches accordingly. For example, the file
system could cache none of the metadatas from the \texttt{untar} phase and run
regime detection during the \texttt{make} phase, resulting in the metadata
server/clients only caching metadata that is repeatedly used. For the job in
Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
(the unit of metadata for file systems) instead of 60K, resulting in almost
40MB (since an inode is about
1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
of memory savings without sacrificing performance.

%- strategy: cache all inodes on create
\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
\caption{The file system metadata requests for a compile job shows distinct
workload phases, characterized by a dominant request type ({\it e.g.}, creates
for ``untar", lookups for ``make", etc.). Using a ParSplice cache management
policy for a workload like this in file systems could reduce memory pressure
without sacrificing performance.  \label{fig:compile-ops}}
\end{figure}

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
automatically size and manage cache tiers
\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain keyspace access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization and without
sacrificing too much performance.
